---
title: 'Project #2'
author: "Kory Martin"
date: "2/26/2023"
output: 
  html_document:
    toc: true
---

## Introduction

For this project, the focus was on gaining continued familiarity with the process of taking untidy data and doing the necessary pre-processing work to convert the data to tidy data. Additionally, we focused on demonstrating our analysis skills on the data, as well as the process of importing data into R and exporting the cleaned up data. 

For this project, I chose the following three datasets that were shared to the class by the following students: 
(a) Coco Donovan - https://raw.githubusercontent.com/rodrigomf5/Tidydata/master/relinc.csv
(b) Alice Ding - https://raw.githubusercontent.com/klmartin1998/cuny_datascience/main/DATA%20607/input/project2_student_data.csv
(c) Kory Martin - https://raw.githubusercontent.com/klmartin1998/cuny_datascience/main/DATA%20607/input/unclean_aatherapist_info.csv

Below, I will demonstrate the work done to import the raw untidy data and then demonstrate the work done to clean the data that will ultimately be used to perform various analysis steps. 

## Setup

We begin the process by loading the libraries that will be used for this project. 


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)

```

## Dataset 1

#### Import and clean data

For this step, we begin by importing the data from a .csv file that is located in GitHub, and then begin doing the work to clean the tidy. The main steps we will take include:

(1) Import the data into a dataframe
(2) Update column names
(3) Convert data from wide to long, by creating a new column labeled "income" and then reassigning the current column values to a new column labeled "num_respondents"


```{r}
ds1 = 'https://raw.githubusercontent.com/rodrigomf5/Tidydata/master/relinc.csv'

df1_parta = read.csv(ds1)

#Initial Data
df1_parta

columns = colnames(df1_parta)
columns

df2_parta <- 
  df1_parta %>%
  rename('under_10k' = columns[2],
         '10k_to_20k' = columns[3],
         '20k_to_30k' = columns[4],
         '30k_to_40k' = columns[5],
         '40k_to_50k' = columns[6],
         '50k_to_70k' = columns[7],
         '75k_to_100k' = columns[8],
         '100k_to_150k' = columns[9],
         'over_150k' = columns[10])

df3_parta <- 
  df2_parta %>% 
  pivot_longer(!religion, names_to = "income", values_to = "num_respondents")

#Cleaned Data
df3_parta

#Output dataframe as .csv file
write_csv(df3_parta, "../output/relinc_tidy.csv")
```

### Analyze Dataset

For this dataset, we performed analyses steps to answer the following questions:

(1) What is the breakdown of all respondents by religious group?
(2) Which religious group has the wealthiest followers?
(3) How do the income breakdowns compare for each religious group?
(4) What percent of respondents refused to report their income per group?
(5) What was the maximum income group for each religion?

**Answers:**
**(1) See chart below**
**(2) When only including the actual respondents to the survey, the top 3 religious groups having the wealthiest followers were Jewish, with 29% of respondents having income over $150K; Hindu, with 25% of respondents having income over $150k; and Athiest, with 17% of respondents reporting having income over $150k.**
**(3) See table below**
**(4) See chart below**
**(5) See table below**

```{r}

df1 <- df3_parta

#(1) What is the religious group makeup of all respondents?
df1 %>%
  group_by(religion) %>%
  summarize(total_respondents = sum(num_respondents)) %>%
  mutate(pct_total = total_respondents/sum(total_respondents)) %>%
  arrange(desc(pct_total)) %>%
  mutate(religion = factor(religion, levels=religion)) %>% 
  ggplot(aes(x=religion, y=pct_total)) +
  geom_bar(aes(x=religion, y=pct_total, fill=religion), stat='identity') +
  theme(axis.text.x = element_text(angle=90)) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Percent of Total Respondents by Religion", 
       x = "Religion", 
       y = "\n Percent of Total Respondents") +
  geom_text(aes(label = paste0(format(pct_total*100,digits=0),"%")))
  


#(2) Which religious group has the wealthiest followers?
df1 %>% 
  filter(income != 'refused') %>% 
  group_by(religion) %>%
  mutate(group_size = sum(num_respondents),
         pct_group = num_respondents/group_size) %>%
  filter(pct_group == max(pct_group),income == 'over_150k') %>%
  arrange(pct_group) %>%
  mutate(religion = factor(religion, levels=religion)) %>%
  ggplot(aes(x=religion, y=pct_group)) +
  geom_bar(aes(fill=religion), stat='identity') +
  labs(title = "\n Top 3 Religious Groups by Followers Income", 
       subtitle = "Pct of Followers having income over $150K",
       x = "\n Religion",
       y="\n Pct of Groups Respondents") +
  scale_y_continuous(labels = scales::percent)
  


#(3) How do the income breakdowns compare for each religious group?
(df1 %>% 
  filter(income != 'refused') %>% 
  mutate(income = factor(income, levels=unique(df1$income))) %>%  
  group_by(religion) %>%
  mutate(group_size = sum(num_respondents),
         pct_group = num_respondents/group_size) %>%
  select(religion, income, pct_group) %>%
  mutate(pct_group = paste0(format(pct_group*100,digits=2),"%")) %>%
  spread(key=income, value=pct_group))




df1 %>% 
  filter(income != 'refused') %>% 
  mutate(income = factor(income, levels=unique(df1$income))) %>%  
  group_by(religion) %>%
  mutate(group_size = sum(num_respondents),
         pct_group = num_respondents/group_size) %>%
  group_by(religion) %>%
  arrange(income) %>%
  ggplot(aes(x=religion, y=pct_group)) +
  geom_bar(aes(x=religion, y=pct_group, fill=income), position='stack', stat='identity') +
  theme(axis.text.x = element_text(angle=90)) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "\n Distribution of Follower Income Groups by Religion", 
       x = "\n Religion",
       fill = "Income Band") + 
  coord_flip()


#(4) What percent of respondents refused to report their income per group?
df1 %>%
  mutate(refused = ifelse(income == 'refused', "Y", "N")) %>%
  group_by(religion) %>%
  summarize(pct_refused = sum(ifelse(refused == 'Y',num_respondents,0))/sum(num_respondents)) %>%
  arrange(desc(pct_refused)) %>%
  mutate(religion = factor(religion, levels=religion)) %>%
  ggplot(aes(x=religion, y=pct_refused)) +
  geom_bar(aes(x=religion, fill=religion),stat='identity') +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Percent of Respondents Not Reporting Income by Religion", 
       x = "\n Religion", 
       y = "\n Percent Refused") +
  scale_y_continuous(labels = scales::percent)

  

#(5) What was the maximum income group for each religion?

df1 %>%
  filter(income != 'refused') %>%
  group_by(religion) %>%
  mutate(group_responses = sum(num_respondents),
         pct_group = num_respondents/group_responses) %>%
  group_by(religion) %>%
  filter(pct_group == max(pct_group)) %>% 
  select(religion, income, pct_group) %>% 
  ggplot() + 
  geom_bar(aes(x=religion, y=pct_group, fill=income), stat='identity') +
  theme(axis.text.x =  element_text(angle=90)) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(title = "\n Plurarity Income Band per Religious Group", 
       x = "\n Religion", 
       y = "\n Percent of Group", 
       fill = "Income Band")
  

```


## Dataset 2


### Import and clean data

For this dataset, I'm going to clean it by completing the following steps:

(1) Import dataset from GitHub
(2) Split the "sex.and.age" column into two separate columns for sex and age respectively
(3) Remove the text from the "test.number" column to make it only hold the numeric value
(4) Change the column names to remove the period from them
(5) Rename the id field to "student_id"
(6) Convert term.1, term.2, and term.3 columns to a single column labeled "term" and move the value to a new column labeled "score"
(7) Remove the text from the newly created "term" column
(8) Change the values for sex to uppercase letters
(9) Convert sex, term, and test_number to be factors


```{r}
ds2 = 'https://raw.githubusercontent.com/klmartin1998/cuny_datascience/main/DATA%20607/input/project2_student_data.csv'

df1_partb <- read.csv(ds2)

df1_partb

df2_partb <- df1_partb %>% 
  mutate(sex = str_match(sex.and.age, "[a-zA-Z]"),
                 age = str_match(sex.and.age, "\\d+"),
                 test.number = str_match(test.number, "\\d+")) %>% 
  select(-sex.and.age) %>%
  rename(student_id = id)
  


df2_partb <- df2_partb %>% janitor::clean_names()

columns = colnames(df2_partb)

df3_partb <- df2_partb %>%
  pivot_longer(columns[5:7], names_to = "term", values_to = "score")

df4_partb <- df3_partb %>% mutate(term = str_match(term, "\\d+"))

df5_partb <- df4_partb %>%
  mutate(sex = toupper(sex)) %>% 
  mutate(test_number = factor(test_number, order = TRUE, levels = c(1,2)), 
         sex = factor(sex), 
         phone = as.character(phone),
         age = as.numeric(age),
         term = factor(term, order = TRUE, levels = c(1,2,3)))


df5_partb

#Output dataframe as .csv file
write_csv(df5_partb, "../output/project2_student_data_tidy.csv")
```


### Analyze Dataset

For this dataset we will answer the following questions:

(1) What is the overall average test score, as well as the average test for each student, each gender, and each test?
(2) How many students are male vs. female?
(3) What is the average age for the group?

**Answers:**
**(1) The average overall test score is 77.7. See chart below for average test score by test number, student, and sex**
**(2) 60% of students were Female and 40% were male**
**(3) Overall the average age of the students was 12.4, with the average age amongst Females being 13, and the average age amongst Males being 11.5**

```{r}
df2 <- df5_partb



#(1) What is the average test score for each student, sex, and test_number?
(average = mean(df2$score))

df2 %>%
  group_by(name) %>%
  summarize(avg_score = mean(score)) %>%
  ggplot(aes(x=name, y=avg_score)) +
  geom_bar(stat='identity', aes(fill=name)) +
  labs(title = "\n Average Test Score per Student", 
       x = "\n Student", 
       y = "\n Average Test Score",
       fill = "Student") +
  geom_hline(yintercept = average, linetype="dashed")

  df2 %>%
  group_by(test_number) %>%
  summarize(avg_score = mean(score)) %>%
  ggplot(aes(x=test_number, y=avg_score)) + 
  geom_bar(stat='identity', aes(fill=test_number)) + 
  labs(title = "\n Average Test Score per Test", 
       x = "\n Test Number", 
       y = "\n Average Test Score",
       fill = "Test Number") +
  geom_hline(yintercept = average, linetype="dashed")



df2 %>%
  group_by(sex) %>%
  summarize(avg_score = mean(score)) %>%
  ggplot(aes(x=sex, y=avg_score)) + 
  geom_bar(stat='identity', aes(fill=sex)) +
  labs(title = "\n Average Test Score per Sex", 
       x = "\n Sex", 
       y = "\n Average Test Score",
       fill = "Sex") +
  geom_hline(yintercept = average, linetype="dashed")



#(3) How many students are male vs. female?
(df2 %>%
  select(name, sex) %>%
  unique() %>%
  group_by(sex) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)))



#(4) What is the average age for the group?
(df2 %>% 
  select(name, sex, age) %>% 
  distinct() %>% 
  summarize(avg_age = mean(age)))


(df2 %>% 
  select(name, sex, age) %>% 
  distinct() %>% 
  group_by(sex) %>%
  summarize(avg_age = mean(age)))


```

## Part 3 - African American Therapist in Los Angeles, CA

For this part of the project, I started off by scraping this [website] (https://www.psychologytoday.com/us/therapists/ca/los-angeles?category=african-american) to create a .csv file with a list of African American therapist in Los Angeles, CA. To clean and convert the data to a tidy format, I made the following transformations


(1) Import the data from .csv file on GitHub to Dataframe
(2) Loop through the existing dataframe and and make updates for each group of rows that correspond to a single therapist
(3) Create new columns X4 and X5 that are the labels for the data in the rows
(4) Using the new columns, spread the data into X2 and X3 such that their values are now assigned to the labels used in X4 and X5
(5) For column with the therapist name (X1), convert blank strings to NA so we can fill in the blanks
(6) For all columns, fill the blanks in the column with any completed value in the column
(7) Take the first row (which is a completed row) and append it to a blank vector
(8) Rename label for X1 to "name"
(9) Remove excess leading and trailing whitespace characters from title and name fields
(10) Remove newline character and any set of 2+ consecutive spaces with a single space
(11) Remove leading and trailing whitespace from city_state_zip field
(12) Sepate the values in the city_state_zip field into separate fields for city, state, and zip
(13) Convert the online and accepting new patients fields to hold single-character codes
(14) Extract the title from the title field (and discard the licensure names)

```{r}
ds3 = 'https://raw.githubusercontent.com/klmartin1998/cuny_datascience/main/DATA%20607/input/unclean_aatherapist_info.csv'

df1_partc <- read.csv(ds3)
df1_partc

df2_partc = c()

start_index = 1
end_index = start_index + 2

while(start_index <= nrow(df1_partc)-4) {
  entry <- df1_partc[start_index:end_index,]
  
  #Create new column titles to correspond with columns X2 and X3
  X4 <- c("title", "statement", "phone")
  X5 <- c("city_state_zip", "online", "new_patients")
  
  #Combine these new columns to the df
  entry <- cbind(entry, X4)
  entry <- cbind(entry, X5)
  
  entry_tib <- as_tibble(entry)
  
  entry_tib <- entry_tib %>% 
    spread(X4,X2) %>%
    spread(X5, X3)
  
  entry_tib <- entry_tib %>% 
    mutate(X1 = ifelse(X1 == "", NA, X1))
  
  entry_tib <- entry_tib %>%
    fill(colnames(entry_tib), .direction="downup")
  
  df2_partc <- rbind(df2_partc, entry_tib[1,])
  
  start_index = end_index+2
  end_index = start_index + 2
  #print(entry)
  
  
}



#Rename column X1 to name
df2_partc <- df2_partc %>% 
  rename(name = X1)

#Strip white space and newline characters from name, statement and title fields

df3_partc <- df2_partc %>% 
  mutate(name = str_squish(name), 
         title = str_squish(title))

#Remove new line and double spacing from the statement field
df3_partc <- df3_partc %>% 
  mutate(statement = str_replace_all(statement, "\\n", "")) %>% 
  mutate(statement = str_replace_all(statement, "\\s{2,}", " "))

#Remove starting and trailing white space from city_state_zip
df4_partc <- df3_partc %>% 
  mutate(city_state_zip = str_squish(city_state_zip))

#Separate the city_state_zip field
df5_partc <- df4_partc %>% 
  separate(city_state_zip, into=c("city", "state_zip"), sep=",") %>% 
  mutate(state_zip = str_squish(state_zip)) %>% 
  separate(state_zip, into=c("state", "zip"))

#Convert the new_patient and online values to single character codes
df6_partc <- df5_partc %>% 
  mutate(new_patients = case_when(new_patients == "Not accepting new clients" ~ "N", new_patients == "Waitlist for new clients" ~ "W", TRUE ~ "Y")) %>% 
  mutate(online = case_when(online == "& Online" ~ "Y", TRUE ~ "N"))



#Simplify the title
df7_partc <- df6_partc %>% 
 mutate(title = str_match(title, pattern=".+?(?=,)")[1])



df7_partc

#Output Tidy data
#write_csv(df7_partc, "../output/aatherapist_tidy.csv")
```

### Analysis

For this dataset we will answer the following:

(1) Quantify the split between Male and Female therapist
(2) Looking at the different types of certifications represented by the population
(3) Identifying the % of therapist offering online, on-site and/or both types of services
(4) Identifying the percent accepting new patients
(5) Break down the group based on their zip code

**Answers**
**(1) The dataset did not have the gender of the therapist, so I was unable to complete this analysis at this time**
**(2) Overall there were 25 distinct title (24 when excluding NA). The top 10 titles represented 96% of the therapists in the group**
**(3) Overall 87.3% of therapist are offering online therapy services**
**(4) Overall 85.36% of therapist were accepting new patients, 9.76% were not accepting new patients, and 4.88%**
**(5) Overall there were 76 unique zip codes represented in the dataset. The top 10 zip codes represented 45.9% of the therapists in the group**

```{r}

length(unique(df3$zip))

df3 <- df7_partc 

df3
#(1) Quantify the split between Male and Female therapist


#(2) Looking at the different types of certifications represented by the population
(sum(df3 %>%
  filter(!is.na(title)) %>%
  group_by(title) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  arrange(desc(n)) %>%
  head(10) %>% 
  select(pct_total)))


df3 %>%
  filter(!is.na(title)) %>%
  group_by(title) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  arrange(desc(n)) %>%
  head(10) %>% 
  select(title, pct_total) %>%
  mutate(pct_total = paste0(format(pct_total*100,digits=1),"%")) %>%
  rename(Title = title, `Percent Total` = pct_total) %>%
  knitr::kable()
  

df3 %>%
  filter(!is.na(title)) %>%
  group_by(title) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  mutate(title = factor(title, levels=title)) %>%
  ggplot(aes(x=title, y=pct_total)) +
  geom_bar(stat='identity', aes(fill=title)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "\n Top 10 Certification Types", 
       x = "\n Title", 
       y = "\n Percent of Therapist", 
       fill = "Certification") + 
  scale_y_continuous(labels = scales::percent)

#(3) Identifying the % of therapist offering online services
(prop.table(table(df3$online)))


df3 %>%
  group_by(online) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  ggplot(aes(x=online, y=pct_total)) +
  geom_bar(aes(fill = online), stat='identity') +
  labs(title = "\n Percent Offering Online Service", 
       y = "\n Percent of Total") +
  scale_y_continuous(labels = scales::percent)


#(4) Identifying the percent accepting new patients
(prop.table(table(df3$new_patients)))

#(5) Break down the group based on their zip code
(sum(df3 %>% 
  group_by(zip) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  select(pct_total)))



df3 %>% 
  group_by(zip) %>%
  summarize(n = n()) %>%
  mutate(pct_total = n/sum(n)) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  mutate(zip = factor(zip, levels=zip)) %>%
  ggplot(aes(x=zip, y=pct_total)) +
  geom_bar(aes(fill=zip), stat='identity') +
  scale_y_continuous(labels = scales::percent) + 
  labs(title = "\n Top 10 Zip Codes", 
       x= element_blank(), 
       y= "Percent of Total") +
  theme(axis.text.x =  element_text(angle=90))

unique(df3$title)

```

## Conclusion


