---
title: 'Assignment #5'
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
date: '2023-02-26'
---
## Introduction

This homework assignment provided applied problems for using the various probability distributions. There were a number of problems that I wasn't too certain of how to answer effectively.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

UNABLE TO FIGURE THIS ONE OUT
```{r}

```

## Question 2

(a) The probability of receiving exactly two inspections after 24-months is **22.3%**
(b) The probability of of receiving 2 or more inspections after 24-months is **33.9%**
(c) The probability of receiving less than 2 inspections after 24-months is **66.1%**
(d) The expected number of inspections is **1.2**
(e) The standard deviation of inspections is **1.07**

```{r}

p = .05
q = 1-p
k = 24

#Probability of receiving exactly 2 inspections after 24-months
dbinom(2, size=k, p)

#Probability of receiving 2 or more inspections after 24-months
pbinom(1,size=k, p, lower.tail=FALSE)

#Probability of receiving fewer than 2 inspections after 24-months
pbinom(1,size=k, p)

#What is the expected number of inspections you should have received?
ev = k*p
print(ev)

#What is the standard deviation?
sd = sqrt(k*p*(1-p))
print(sd)
```
## Question 3 

(a) The probability that exactly 3 patients arrive in one hour is **.76%**
(b) The probability that more than 10 patients arrive in one hour is **42%**
(c) In 8 hours, we would expect **80 patients** to arrive
(d) The standard deviation of this probability distribution is **3.16**
(e) DON'T UNDERSTAND THE QUESTION

```{r}
lambda = 10

#What is probability that exactly 3 arrive in one hour?
k=3

dpois(k,lambda)

((lambda^k)*(exp(-lambda)))/factorial(k)

#What is the probability that more than 10 arrive in one hour?
x = 10
ppois(x, lambda, lower.tail = FALSE)

#How many would you expect to arrive in 8 hours?


#What is the standard deviation of the appropriate probability distribution?
sd = sqrt(lambda)
print(sd)
```

## Question 4

For this problem, there are 30 total employees, with 15 of these being nurses and 15 non-nurses. This results in a probability of selecting a nurse in any single event (p) of 50%; and the probability of selecting a non-nurse (q) of 50%.

(a) The probability of selective 5 nurses out of 6 trips is **7.6%**
(b) The expected number of nurses in 6-trips is **3**
(c) The expected number of non-nurses in 6-trips is **3**

```{r}
k = 
p = 15/30
q = 1-p

#What is the probability of selecting five nurses out of 6 trips?
M = 15
N = 30
n = 6
x = 5

(choose(M,x)*choose(N-M,n-x))/choose(N,n)

#What is the expected number of nurses for 6 trips
6 * p

#How many non-nurses would we have expected subordinate to send?

```

## Question 5

(a) The probability the driver will be seriously injured during the course of 12 months is **67%**
(b) The probability the driver will be seriously injured during the course of 15 months is **77.7%**
(c) The expected number of hours that a driver drives before he is seriously injured is **1000 hours**
(d) The probability that the driver will have a serious injury in next 100-hours given no injury in 1200 hours is **72.7%**

```{r}


lambda = .001
lambda * 1200
q = 1-p


#What is probability that the driver will be seriously injured during course of 12-months?
n=0
p0 = (exp(-1.2)*(1.2^n))/factorial(n)
1-p0



#What is the probability that driver will be seriously injured during course of 15-months?

k = 1200
num_months = 15
hours_monthly = k/12

total_hours = num_months*hours_monthly
lambda = .001*total_hours


n=0
p0 = (exp(-lambda)*(lambda^n))/factorial(n)
1-p0



#What is expected number of hours that a driver will drive before being seriously injured?
ev = 1/.001


#Given that a driver has driven 1200 hours, what is probability that they will be injured in the next 100 hours?


#Probability of injury in 100-hours is
n=0
p0 = (exp(-1.3)*(1.3^n))/factorial(n)
pa = 1-p0
pb =  (exp(-1.2)*(1.2^n))/factorial(n)
pab = pa*pb
cond_prob = pab/pb



```

## Question 6

(a) The probability the generator will fail more than twice in 1000 hours is **8%**
(b) The expected value is **1**

```{r}
lambda = 1
p = 1/1000
q = 1-p

#What is the probability the generator will fail more than twice in 1000 hours?
p0 = (exp(-1)*lambda^0)/factorial(0)
p1 = (exp(-1)*lambda^1)/factorial(1)
p2 = (exp(-1)*lambda^2)/factorial(2)

p = p0+p1+p2
1-p


#What is the expected value?


```

## Question 7

(a) The probability the patient will wait more than 10-minutes is **66%**
(b) The probability the patient will wait at least another 5 minutes after waiting 10-minutes is **55%**
(c) The expected wait time is **15-minutes**

```{r}

#What is the probability the patient will wait more than 10-minutes?
1- punif(10,0,30)

#If patient has already waited 10 minutes, what is probability they will wait another 5 minutes?
1-punif(5,0,30)



#What is expected waiting time?
a = 0
b = 30
ev = (a+b)/2
ev

```

## Question 8

(a) The expected failure time is **.10**
(b) The standard deviation is also **.10**
(c) The probability of failure after 8-years is **45%**
(d) The probability of failure in next 2-years after already owning it for 8-years is **8.1%**

```{r}

#What is expected failure time?

ev = 1/10

#What is the standard deviation?

sd = sqrt(1/100) 

#What is probability of failure after 8-years?
1-pexp(8,ev)


#What is probability of failure in 2-years given you already owned it for 8-years?
p2 = pexp(2,ev)



```